---
title: "bowling"
author: "Lewis Pipkin"
date: "January 12, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(33)
library(tidyverse)
```

Last night, my good friend and I went bowling- $36 for 100 minutes of bowling (which would be a fantastic deal if it included shoe rental) and a prime opportunity for data collection. I lost 5 of 6 games, but most of the games were tantalizingly close. So, I ran a Monte Carlo simulation to see what my record would be after 100k matchups (not taking alcohol consumption into account- I'm a teetotaler but my friend bought some Dos Equis at the bowling alley's concession stand). It wasn't pretty:


```{r simulation}
Lewis <- c(90,79,90,102,94,109)
Erik <- c(139,97,104,110,99,100)

#mean(Lewis) #94
#sd(Lewis) #10.44988

#mean(Erik) #108.1667
#sd(Erik) #15.79135

lew <- c()
er <- c()
for(i in 1:1e5){
  lew[i] <- rnorm(1,mean(Lewis),sd(Lewis))
  er[i] <- rnorm(1,mean(Erik),sd(Erik))
}
df <- data.frame(lew,er)
df$winner <- ifelse(df$lew > df$er, "Lewis", "Erik")
sum(df$winner=="Lewis")
sum(df$winner=="Erik")
```

Sheesh. Have you ever lost 77000 of something? It doesn't feel too good. But, how did we get here? Let's dive into the data:

```{r}
bowling <- read_csv("~/Documents/bowling.csv")
head(bowling)
bowling %>% group_by(Name) %>% summarize(avg_pins=mean(pins))
```
So, I had a lower pins-per-roll, naturally. Here is the density:

```{r}
ggplot(bowling,aes(x=pins, fill=Name)) + geom_density(alpha=0.25)
```

And we ended up being pretty even in terms of spares and strikes.

```{r}
bowling %>% group_by(Name) %>% summarize(tot_spare=sum(Is_spare),
                                         tot_strike=sum(Is_strike))
```

Let's see how many pins were left standing going into each of these 25 total spares:

```{r}
table(10-bowling$pins[which(bowling$Is_spare==1)-1])
```

So, 3 of these were essentially strikes, but the bulk of them were with 3 or fewer pins left going into the roll. Makes sense. 

But now, we'll check out how confidence and repetion played into these matchups. After a spare or a strike, average number of pins shot up (compare this to the first table):

```{r}
bowling_ <- bowling[which(bowling$Is_spare==1 | bowling$Is_strike==1)+1,]
bowling_ %>% group_by(Name) %>% summarize(avg_pins=mean(pins))
```

And here is our average pins per frame over these 6 games:

```{r}
b <- bowling %>% group_by(Name, Frame) %>% summarize(avg_pins=mean(pins))
ggplot(b,aes(x=Frame,y=avg_pins,col=factor(Name))) + geom_point() + geom_line()
```

As we can see, I started off stronger, but soon enough I would begin to taper off. (Or, rather, Erik would begin to pull away.) We see that below, in terms of average score per frame, incorporating the post-spare or -strike bonuses:


```{r}
b_ <- bowling %>% group_by(Name, Frame) %>% summarize(avg_score=mean(cumul_score))
ggplot(b_,aes(x=Frame,y=avg_score,col=factor(Name))) + geom_point() + geom_line()

```

Oftentimes I was close behind if not leading halfway through, and then I wouldn't be able to put him away. I realize that since I was already owned extremely hard last night in 5 of the 6 games, I didn't have to write this and prove to the world that I am bad at bowling with tables and charts, but I did it anyway-- for the love of knowledge and discovery. I'll probably throw up the bumpers next time, though.








